---
title: "Lab 1 - TDDE07"
author: "Axel Holmberg (axeho681), Wilhelm Hansson (wilha431)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## 1.

### a)

```{r}
#1 a)
set.seed(12345)
library(ggplot2)
a0 = 2
b0 = 2
s = 5
f = 15
x_vector = seq(1, 10000, 1)
set.seed(12345)
distribution =rbeta(x_vector, a0+s, b0+f)

y_calced_mean = c(1:length(x_vector))
for (i  in x_vector) {
	y_calced_mean[i] = mean(distribution[0:i])
}

qplot(x=x_vector,y=y_calced_mean, geom="line")+
	labs(x="Number of draws", y="Mean", title="Mean as a function of number of draws")

```

Above is a graph that shows the mean as a function of the number of draws. As one can see it converges to about $0.292$. To calculate the true value for the mean of the posterior the following values are used $\alpha =\alpha_o + s  = 2  + 5$ and $\beta = \beta_0 + f = 2 + 15$ and the formula to calculate the mean of a beta distribution is $E(\theta|y) = \alpha / (\alpha + \beta)$ where $\theta|y \sim Beta(a_0 + s, b_0+f)=Beta(7,17)$. The true value of the mean is $0.2917$.

```{r}

y_calced_sd = c(1:length(x_vector))
for (i  in x_vector) {
	y_calced_sd[i] = sd(distribution[0:i])
}

qplot(x=x_vector,y=y_calced_sd, geom="line")+
	labs(x="Number of draws", y="Standard deviation", title="Standard deviation as a function of number of draws")


```

Above is a graph that shows the standard deviation as a function of the number of draws. As one can see it converges to about $0.091$. To calculate the true value for the standard deviation of the posterior the following values are used $\alpha =\alpha_o + s  = 2  + 5$ and $\beta = \beta_0 + f = 2 + 15$ and the formula to calculate the standard deviation of a beta distribution is $SD(\theta|y)= \sqrt{V(\theta|y)} = \sqrt{\alpha \beta / ((\alpha + \beta)^2(\alpha + \beta + 1))}$ where $\theta|y \sim Beta(a_0 + s, b_0+f)=Beta(7,17)$. The true value of the standard deviation is $0.0901$.

### b)

```{r}
#1 b)

classified = ifelse(distribution > 0.3, 1, 0)
classified_table = table(classified)
counted_classified = classified_table[names(classified_table)==1]

classified_probabilty = counted_classified/length(distribution)
classified_probabilty

true_probability_value = 1- pbeta(0.3, a0+s, b0+f)

```

The true value of the probability is `r true_probability_value` and the the calculated probability is `r classified_probabilty`. They are almost equal.

### c)

```{r}
#1 c)

log_odds = log(distribution/(1-distribution))
hist(log_odds)

density(log_odds)

```

Above is a histogram of the log-odds. The log-odds seems to be normally distributed with a mean of about $-1$. 

## 2.

### a)

```{r}
#2 a)

library(LaplacesDemon)
observations <- c(44, 25, 45, 52, 30, 63, 19, 50, 34, 69)
n <- length(observations)
my <- 3.7
number_of_draws <- 10000
set.seed(12345)


tausquared <-
	sum((log(observations) - my) ^ 2) / length(observations)

sigma_sq <- rinvchisq(number_of_draws, n, tausquared)

interval = seq(min(sigma_sq), max(sigma_sq), 0.001)
invchisq = dinvchisq(interval, n, tausquared)


plot <-
	ggplot() + geom_histogram(aes(x = sigma_sq, y = ..density..), fill = "gray", binwidth = 0.02) + geom_line(aes(x =interval, y = invchisq), position = "stack") + labs(x="Sigma Squared", y = "Density", title="Draws from posterior against the theoretical posterior distributon.")

plot

```

Above is the graph for the draws from posterior plotted against the theoretical posterior distribution. What this shows is that the draws from the posterior fits the theoretical posterior distribution well.

### b)

```{r}
#2 b)

gini <- 2*pnorm(sqrt(sigma_sq)/sqrt(2))

plot_gini <- ggplot() + geom_histogram(aes(x=gini, y=..density..), binwidth = 0.001)
plot_gini

```

Plot show density of calculated gini coefficients. The coefficients are calculated from the draws of a).

### c)

```{r}
#2 c)

library(HDInterval)

interval_pin <- p.interval(gini, HPD=FALSE, prob=0.9)

gini_density <- density(gini)

interval_hdi <- hdi(gini_density, credMass=0.9)



plot_CI <- plot_gini + geom_vline(aes(xintercept=c(interval_pin[1,1], interval_pin[1,2]),color="ETI")) + 
	geom_vline(aes(xintercept=c(interval_hdi["lower"], interval_hdi["upper"]),color="HDI")) + scale_color_manual(name = '',
    values = c(
      'ETI' = 'red',
      'HDI' = 'green'
    ))
plot_CI

```

In the graph above both the ETI and the HDI interval. Where the HDI-interval is a CI with respect to the density and the ETI just takes the CI of the values without respect to the density. If one has a non-symmetrical distribution, one should use a HDI rather than an ETI when calculating CI as it takes the density into consideration.

## 3.

### a)

```{r}
#3 a)

directions <-
	c(-2.44, 2.14, 2.54, 1.83, 2.02, 2.33, -2.79, 2.23, 2.07, 2.02)

mu <- 2.39

p_y_given_k_mu <- c()

k_v = seq(0.001,10,0.01)

posterior <- c()
likelihood <- c()
prior <- c()

for (o in 1:length(k_v)) {
	probs <- c()
	k <- k_v[o]
	for (i in 1:length(directions)) {
		y <- directions[i]
		
		probs[i] <- exp(k * cos(y - mu)) / (2 * pi * besselI(k, 0)) #likelihood
	}
	likelihood[o] <- prod(probs)
	posterior[o] <- prod(probs) * dexp(k,1) #product of likelihood for y1...yn * prior for k
	prior[o] <- dexp(k,1)
}


plotposterior <-
	ggplot() + geom_line((aes(
		x = k_v,
		y = posterior / sum(posterior),
		color = "Posterior"
	))) + geom_line((
		aes(
			x = k_v,
			y = likelihood / sum(likelihood),
			color = "Likelihood"
		)
	))  + geom_line((aes(
		x = k_v,
		y = prior / sum(prior),
		color = "Prior"
	))) + labs(x = "k", y = "Density", title = "Posterior distribution of k") + scale_color_manual(
		name = '',
		values = c(
			'Prior' = 'red',
			'Posterior' = 'green',
			'Likelihood' = 'blue'
		)
	)
plotposterior

```

The posterior distribution of $\kappa$ is as expected based on the density of the prior and the likelihood.

### b)

```{r}
#3 b)

plotkmode <- plotposterior + geom_vline(aes(xintercept=k_v[which.max(posterior)]))
plotkmode

posterior_mode <- k_v[which.max(posterior)]


```

The mode of the posterior is `r posterior_mode`.

\newpage

## Appendix for code

```{r, code=readLines("Lab_1.R"), echo=TRUE, eval=FALSE}

```